---
title: "exercise-09"
output: html_document
date: "2024-03-18"
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

### Load in Libraries
```{r}
library(tidyverse)
library(skimr)
library(broom)
library(infer)
```

### Load in dataset
```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Street_et_al_2017.csv"

d <- read_csv(f)
```

### Generate 5 number summary, mean, and standard deviation of each
### quantitative variable
```{r}
stats <- d %>% skim() %>% select(, c(skim_variable, numeric.mean, numeric.sd, numeric.p0, numeric.p25, numeric.p50, numeric.p75, numeric.p100)) %>% na.omit() %>% rename(, Variable = skim_variable, Mean = numeric.mean, Standard_Deviation = numeric.sd, Min = numeric.p0, Q1 = numeric.p25, Median = numeric.p50, Q3 = numeric.p75, Max = numeric.p100) %>% as.data.frame()

stats
```

### Plot brain size (ECV) as a function of social group size (Group_size),
### longevity (Longevity), juvenile period length (Weaning), and
### reproductive lifespan (Repro_lifespan)
```{r}
ecvGS <- ggplot(d, aes(Group_size, ECV)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)

ecvLong <- ggplot(d, aes(Longevity, ECV)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)

ecvJPL <- ggplot(d, aes(Weaning, ECV)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)

ecvRLS <- ggplot(d, aes(Repro_lifespan, ECV)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)


ecvGS
ecvLong
ecvJPL
ecvRLS
```

### Calculate the ordinary least squares regression coefficients B1 and 
### B0 for ECV as a function of social group size.
```{r}
sGS <- select(d, c('Species', 'Taxonomic_group', 'ECV', 'Group_size')) %>% na.omit()

x <- sGS$Group_size
y <- sGS$ECV
b1 <- cov(x, y)/var(x)

b0 <- mean(y) - b1 * mean(x)

b1
b0
```

### Confirm B1 and B0 using lm()
```{r}
m <- lm(y ~ x, data = sGS)

m
```

### Repeat analysis for Catarrhines, Platyrrhines, and Strepsirhines
##### The slope coefficients differ slightly from each other, though there
##### is a large difference between the intercepts. 
```{r}
# Catarrhinnes
sC <- subset(sGS, Taxonomic_group == 'Catarrhini')

xC <- sC$Group_size
yC <- sC$ECV
b1C <- cov(xC, yC)/var(xC)

b0C <- mean(yC) - b1C * mean(xC)

mC <- lm(yC ~ xC, data = sC)

# Platyrrhines

sP <- subset(sGS, Taxonomic_group == 'Platyrrhini')

xP <- sP$Group_size
yP <- sP$ECV
b1P <- cov(xP, yP)/var(xP)

b0P <- mean(yP) - b1P * mean(xP)

mP <- lm(yP ~ xP, data = sP)

# Strepsirhines

sS <- subset(sGS, Taxonomic_group == 'Strepsirhini')

xS <- sS$Group_size
yS <- sS$ECV
b1S <- cov(xS, yS)/var(xS)

b0S <- mean(yS) - b1S * mean(xS)

mS <- lm(yS ~ xS, data = sS)

Taxonomic_Group <- c('Catarrhines', 'Platyrrhines', 'Strepsirhines','All')
Slope_Coeff_B1 <- c(b1C, b1P, b1S, b1)
Intercept_B0 <- c(b0C, b0P, b0S, b0)

Tax_Stats <- data.frame(Taxonomic_Group, Slope_Coeff_B1, Intercept_B0)
Tax_Stats
```

### Calculate standard error of the slope coefficient, the 95% confidence
### confidence interval, and the p-value of coefficient
```{r}
### Standard error of slope coefficient

n <- nrow(sGS)

seB1 <- sqrt(((sum(y-mean(y)))^2)/((n-2)*(sum(x-mean(x)))^2))
seB1

### 95% CI
alpha <- 1 - 95/100

lower <- b1 - qt(1 - alpha/2, df = n - 2) * seB1
upper <- b1 + qt(1 - alpha/2, df = n - 2) * seB1
CI <- cbind(lower, upper)
CI

### p-value
tstat <- (b1 - 0)/seB1
pval <- 2 * pt(tstat, df = n - 2, lower.tail = FALSE)
pval
```

### Generate a null sampling distribution for the slope coefficient from
### 1000 permutations. Permutations consist of randomly shuffling the
### response variable from the explanatory variable.
```{r}
perm <- sGS %>%
  specify(ECV ~ Group_size) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "slope")

pvalPerm <- perm %>%
  get_p_value(obs_stat = b1, direction="two_sided")

pvalue <- perm %>% 
  mutate(abs_stat=abs(stat)) %>%
  summarize(
    estimate = mean(abs_stat >= abs(b1))
  )

pvalPerm
pvalue
```

### Generate a 95% CI of the slope coefficient estimate using
### bootstrapping and both the percentile method and the theory-based
### method
##### The slope is significantly different from zero because the CI does
##### not contain zero
```{r}
boot_slope_dist <- sGS %>%
  specify(ECV ~ Group_size) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "slope")

slope_estimate <- mean(boot_slope_dist$stat)

percentCI <- get_ci(boot_slope_dist, level = 1 - alpha, type = "percentile")
theoryCI <- get_ci(boot_slope_dist, level = 1 - alpha, type = "se", point_estimate = slope_estimate)

```



